---
title: "Accessible colour assessment"
author: "NDRS"
date: "`r Sys.Date()`"
output: html_document
params:
  palette: c("#12436D","#28A197","#801650","#F46A25","#3D3D3D","#A285D1")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(DT)
library(colorspace)
library(dichromat)
library(recolorize)
```

If wanting to create accessible visulisations please take the time to read through the [Government Analysis function guide to data vis](https://analysisfunction.civilservice.gov.uk/policy-store/data-visualisation-colours-in-charts/) and try to use as much of the advice as possible. To this end this tool has been created to assess your colour palette in line with advice given by the Government Analysis function (AF).
## Palette Visualisation

```{r, echo=FALSE}
palette <- params$palette

contrast_matrix <- function(x, palette) {
  round(colorspace::contrast_ratio(col = x, col2 = palette),2)
}

styleTable <- function(data){
    datatable(data, 
              options = list(
                  rowCallback = JS(
                      "function(row, data) {",
                      "for(var i = 0; i < data.length; i++) {", 
                      "if(data[i] > 3) {",
                      "$('td:eq('+i+')', row).css({'background-color': '#A9CDBC'});",
                      "}", 
                      "}", 
                      "}")
              )
    )
}

scales::show_col(palette)
```

## Contrast to white

The AF advises that all colours should have contrast ratio of over 3 compared to white background.


```{r, echo=FALSE}

contrast_res <- sapply("white", contrast_matrix, palette = palette)

rownames(contrast_res) <- palette

styleTable(contrast_res)

```

## Contrast within palette

It is advised that all adjacent colours achieve a contrast ratio of 3. 

Ideally a plot should not show anymore than 4 colours.

If using hues of a single colour for sequential data be aware that the mid colour will not achieve the required ratio of contrast to the dark and light colours.


```{r, echo=FALSE}



contrast_res <- sapply(palette, contrast_matrix, palette = palette)

rownames(contrast_res) <- colnames(contrast_res)

styleTable(contrast_res)

```

## Consider how your colour palette looks to people who are colour blind

There are 3 main types of colour blindness. Below is a best estimation of what colour blind people are likely to see when viewing your colour palette.

```{r, echo=FALSE}
protan <- dichromat(palette, type = "protan")
deutan <- dichromat(palette, type = "deutan")
tritan <- dichromat(palette, type = "tritan")

# plot for comparison
layout(matrix(1:4, nrow = 4)); par(mar = rep(1, 4))
plotColorPalette(palette, main = "Trichromacy")
plotColorPalette(protan, main = "Protanopia")
plotColorPalette(deutan, main = "Deutanopia")
plotColorPalette(tritan, main = "Tritanopia")
```
